The Radio Sky at Meter Wavelengths: $m$-Mode Analysis Imaging\\ with the Owens Valley Long Wavelength Array

### 2017-09-21 paper.tex
fix some typos and compute B*v for the rfi

Current 21~cm cosmology experiments can be broadly separated into two classes: global experiments
that are aiming to detect the spectral signature of the cosmologically redshifted 21~cm transition
after averaging over the entire sky, and power spectrum experiments that incorporate angular
information to attempt to measure the 3D spatial power spectrum of cosmological 21~cm perturbations.
Ongoing global experiments include EDGES \citep{2010Natur.468..796B, 2017ApJ...835...49M}, LEDA
\citep{todo_price_2017}, BIGHORNS \citep{2015PASA...32....4S}, SCI-HI \citep{2014ApJ...782L...9V},
and SARAS 2 \citep{2017arXiv170306647S}.  Ongoing power spectrum experiments include PAPER/HERA
\citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR \citep{2017ApJ...838...65P}, and the MWA
\citep{2016ApJ...833..102B, 2016MNRAS.460.4320E}.

@@

This paper presents a series of new low-frequency maps of the sky between 36.528~Mhz and 73.152~Mhz,
capturing the full sky visible from the Owens Valley Radio Observatory (OVRO) with angular
resolution of $\sim 15$~arcmin. We aim for these maps to inform foreground removal strategies in
21~cm cosmology as they represent some of the highest angular resolution full-sky maps ever produced
at these frequencies. The maps will be made freely available online at the Legacy Archive for
Microwave Background Data Analysis (LAMBDA)\footnote{\url{https://lambda.gsfc.nasa.gov/}}.

@@

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/terrestrial-interference/smeared}
    \caption{
        This figure illustrates terrestrial sources of correlated noise that are apparent after
        averaging the visibilities at 62.688 MHz over the entire 28 hour observing period (keeping
        the phase center at zenith such that astronomical sources of radio emission are smeared
        along tracks of constant declination). Each panel represents a different component that is
        removed from the visibilities. The images are generated using WSCLEAN
        \citep{2014MNRAS.444..606O}, uniform weighting, and only baselines longer than 15
        wavelengths. Panels (a) and (b) illustrate components that appear noise-like in image-space,
        but are in fact a constant-offset to the measured visibilities likely associated with
        cross-talk or common-mode pickup. Panel (c) illustrates a component that is clearly
        associated with an RFI source on the horizon to the west-north-west of the OVRO-LWA. This
        RFI source is likely an arcing power line.  Figure~\ref{fig:rings} illustrates the
        characteristic ring-like artifacts introduced into the maps if these 3 components are not
        removed prior to $m$-mode analysis imaging.
    }
    \label{fig:fitrfi}
\end{figure*}

\begin{figure*}[t]
    \centering
    \includegraphics[height=0.32\textheight]{figures/terrestrial-interference/rings}
    \caption{
        This is a Mollweide-projected image of the artifacts introduced to the $m$-mode analysis
        maps by the three terrestrial sources shown in Figure~\ref{fig:rings}. Because these sources
        are not moving through the sky sidereally, they tend to be smeared along rings of constant
        declination. The spurs seemingly radiating from the north celestial pole are a Moir\'{e}
        pattern (ie. an artifact of the pixelization).
    }
    \label{fig:rings}
\end{figure*}

@@

For instance, a bright interfering source might contribute $\b v_\text{terrestrial}$ to the measured
$m$-modes. However, if $\b v_\text{terrestrial}$ is actually perpendicular to all of the columns of
$\b B$, there will be no contamination in the map because $\b B^*\b v_\text{terrestrial} = \b 0$.
In practice this is unlikely. In general the contamination is proportional to the overall amplitude
of the interference ($\|\b v_\text{terrestrial}\|$) and the degree to which the interference mimics
an astronomical signal ($\|\b B^*\b v_\text{terrestrial}\|/\|\b v_\text{terrestrial}\|$).

\todo{Compute B*v for Figure~\ref{fig:fitrfi} to show that the pick-up is pretty bad. My expectation
is that it is actually worse than the RFI, but I don't actually know for sure}

>>>>

Current 21~cm cosmology experiments can be broadly separated into two classes: global experiments
that are aiming to detect the spectral signature of the cosmologically redshifted 21~cm transition
after averaging over the entire sky (otherwise known as the monopole), and power spectrum
experiments that incorporate angular information to attempt to measure the 3D spatial power spectrum
of cosmological 21~cm perturbations.  Ongoing global experiments include EDGES
\citep{2010Natur.468..796B, 2017ApJ...835...49M}, LEDA \citep{todo_price_2017}, BIGHORNS
\citep{2015PASA...32....4S}, SCI-HI \citep{2014ApJ...782L...9V}, and SARAS 2
\citep{2017arXiv170306647S}.  Ongoing power spectrum experiments include PAPER/HERA
\citep{2015ApJ...809...61A, 2016arXiv160607473D}, LOFAR \citep{2017ApJ...838...65P}, and the MWA
\citep{2016ApJ...833..102B, 2016MNRAS.460.4320E}.

@@

This paper presents a series of new low-frequency maps of the sky between 36.528~MHz and 73.152~MHz,
capturing the full sky visible from the Owens Valley Radio Observatory (OVRO) with angular
resolution of $\sim 15$~arcmin. We aim for these maps to inform foreground removal strategies in
21~cm cosmology as they represent some of the highest angular resolution full-sky maps ever produced
at these frequencies. The maps will be made freely available online at the Legacy Archive for
Microwave Background Data Analysis (LAMBDA)\footnote{\url{https://lambda.gsfc.nasa.gov/}}.

@@

\begin{figure*}[t]
    \includegraphics[width=\textwidth]{figures/terrestrial-interference/smeared}
    \caption{
        This figure illustrates terrestrial sources of correlated noise that are apparent after
        averaging the visibilities at 62.688 MHz over the entire 28 hour observing period (keeping
        the phase center at zenith such that astronomical sources of radio emission are smeared
        along tracks of constant declination). Each panel represents a different component that is
        removed from the visibilities. The images are generated using WSCLEAN
        \citep{2014MNRAS.444..606O}, uniform weighting, and only baselines longer than 15
        wavelengths. Panels (a) and (b) illustrate components that appear noise-like in image-space,
        but are in fact a constant-offset to the measured visibilities likely associated with
        cross-talk or common-mode pickup. Panel (c) illustrates a component that is clearly
        associated with an RFI source on the horizon to the west-north-west of the OVRO-LWA. This
        RFI source is likely an arcing power line.  Figure~\ref{fig:rings} illustrates the
        characteristic ring-like artifacts introduced into the maps if these 3 components are not
        removed prior to $m$-mode analysis imaging. The component shown in panel (a) has about twice
        the amplitude ($\|\b v_\text{terrestrial}\|$) of panels (b) and (c), and for all three
        components $\|\b B^*\b v_\text{terrestrial}\|/(\|\b B\|\|\b v_\text{terrestrial}\|) \sim
        0.035$.
    }
    \label{fig:fitrfi}
\end{figure*}

\begin{figure*}[t]
    \centering
    \includegraphics[height=0.32\textheight]{figures/terrestrial-interference/rings}
    \caption{
        This is a Mollweide-projected image of the artifacts introduced to the $m$-mode analysis
        maps by the three terrestrial sources shown in Figure~\ref{fig:fitrfi}. Because these
        sources are not moving through the sky sidereally, they tend to be smeared along rings of
        constant declination. The spurs seemingly radiating from the north celestial pole are a
        Moir\'{e} pattern (ie. an artifact of the pixelization).
    }
    \label{fig:rings}
\end{figure*}

@@

For instance, a bright interfering source might contribute $\b v_\text{terrestrial}$ to the measured
$m$-modes. However, if $\b v_\text{terrestrial}$ is actually perpendicular to all of the columns of
$\b B$, there will be no contamination in the map because $\b B^*\b v_\text{terrestrial} = \b 0$.
In practice this is unlikely. In general the contamination is proportional to the overall amplitude
of the interference ($\|\b v_\text{terrestrial}\|$) and the degree to which the interference mimics
an astronomical signal ($\|\b B^*\b v_\text{terrestrial}\|/(\|\b B\|\|\b v_\text{terrestrial}\|)$).
